\documentclass[12pt]{article}
\usepackage{geometry}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{parskip}
\usepackage{enumerate}
\usepackage{stmaryrd}
\usepackage{listings}
\usepackage{fullpage}

\begin{document}

\title{CS 350 Notes}
\author{Matthew Visser}
\date{Sep 22, 2011}
\maketitle

\section{Mutual Exclusion}

\begin{itemize}
    \item A thread can be preempted at any time in Peterson's algorithm.
    \item Peterson's algorithm only works for two threads. 
\end{itemize}

\subsection{Starvation}

A thread may never get to execute because of the way that the synchronization
algorithm or scheduler works.

Can use special instructions:
\begin{itemize}
    \item Test-and-set
    \item Swap
\end{itemize}

\subsection{Test-and-Set}

Sets the value of a memory location and applies a condition on the old value.
Since this is atomic, only one thread will execute it first, leaving the others
with the new value.

\subsection{Swap}

Can swap and then test old value.

\subsection{How do we wait until it's our turn?}

You can either spin or block. Spinning is busy waiting, such as
\texttt{while( condition ) ;}. Blocking is sleeping. If you don't expect to spin
for long, then it may be better to spin than yield. Otherwise, if it will take a
long time, it may be better to yeild.

\section{Other Synchronization Primitives}

\subsection{Semaphores}

\begin{itemize}
    \item Can be used to enforce mutual exclusion requirements. It can also
        enforce other synchronization problems.
    \item Has integer value and supports two operations:
        \begin{itemize}
            \item P $\to$ Wait --- call this before starting the critical
                section.
            \item V $\to$ Signal --- call this after leaving the critical
                section.
        \end{itemize}
    \item Can allow multiple threads to access a critical section, but limit the
        number.
    \item Need to enforce that \texttt{P()} and \texttt{V()} are atomic. If on a multi-core CPU,
        \texttt{P()} \texttt{V()} themselves are critical sections.
\end{itemize}

\subsection{Thread Blocking in OS161}
\begin{itemize}
    \item \texttt{thread\_sleep(const void *addr)} --- sleeps the current thread.
    \item \texttt{thread\_wakeup(const void *addr)} --- wakes up \underline{all}
        threads that are sleeping on \texttt{addr}.
\end{itemize}


\end{document}
% vim: tw=80
